---
title: "Analysis: Construction Repetition"
output: html_document
date: '2022-05-19'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lmerTest)
library(lme4)
# library(tidyverse)
library(ggplot2)
library(readr)
library(magrittr)
library(dplyr)
#
library(reshape2)
# library(rjags)
# library(BayesianFirstAid)
library(effects)
library(jtools)
```


## Load data and rename variables for convenience
```{r,  echo=FALSE}
data <- read_csv('/Users/mario/code/exp-rep/paper/aacl2022/submission/code/chains_all_SBNC_gpt2_50_1e-3.csv')
# data$logFE1 <- log2(data$FE1)
# data$logFE2 <- log2(data$FE2)
# data$logFE3 <- log2(data$FE3)
# data$logFE4 <- log2(data$FE4)
# data$logFE5 <- log2(data$FE5)
# data$logFE10 <- log2(data$FE10)
# data$logFE15 <- log2(data$FE15)
# data$logFE20 <- log2(data$FE20)
# data$logFE25 <- log2(data$FE25)
# data$logFE30 <- log2(data$FE30)
data$`Facilitating Effect` <- data$FE
data$CurrentTurn <- data$CurrentTurn + 1
data$`log Utterance Position` <- log2(data$CurrentTurn)
data$`log Information Content` <- log2(data$S)
data$`log Distance` <- log2(data$RecencyBoth)
# data$logRecencyOther <- log2(data$`RecencyOther`)
# data$logRecencySame <- log2(data$`RecencySame`)
data$RecencyOther <- NULL
data$RecencySame <- NULL
data$SS <- NULL
data$`log Repetition Index` <- log2(data$`RepetitionIndex`)
data$`Repetition Index` <- data$RepetitionIndex - 1
data$logCurrentSpan <- log2(data$`CurrentSpanningTokens`)
data$logTotalSpan <- log2(data$`TotalSpanningTurns`)
data$`log Construction Length` <- log2(data$`Length`)
data$logFreq <- log2(data$`Frequency`)
data$logFreeFreq <- log2(data$`FreeFrequency`)
data$`log Repetition Index in Utterance` <- log2(data$`IndexInTurn` + 1)
data$`Previous Same Utterance` <- data$PreviousInSameTurn
data$`Referential` <- data$Topical
```

### First mentions vs. repetitions

Split usages according to whether they are first mentions (RepetitionIndex = 1) or repetitions (RepetitionIndex > 1).

```{r}
tags <- c("\nFirst mentions", "\nRepetitions")

dH1 <- as_tibble(data) %>% 
   mutate(Repetition = case_when(
     `Repetition Index` == 0 ~ tags[1],
     `Repetition Index` >= 1  ~ tags[2]
))

dH1$Repetition <- factor(dH1$Repetition, levels = tags, ordered = FALSE)
summary(dH1$Repetition)
```


Box plot of facilitating effect for first mentions and repetitions

```{r message=FALSE, warning=FALSE}
pdf(file = "/Users/mario/code/exp-rep/paper/aacl2022/figures/fe-rep-first.pdf",
    width = 6,
    height = 6)

ggplot(data = dH1, mapping = aes(x=Repetition,y=FE)) + 
  geom_jitter(aes(color='blue'),alpha=0.2) +
  geom_boxplot(fill="bisque",color="black",alpha=0.3,range=0.1,outline=FALSE) + 
  labs(x='') + labs(y='') +
  guides(color=FALSE) +
  theme_minimal() +
  theme(text = element_text(size=25))
dev.off()
```

#### T-tests
```{r}
t.test(FE ~ Repetition, data=dH1)
t.test(S ~ Repetition, data=dH1)
```


### Cumulativity 

Group repetition indices into bins
```{r}
tags <- c("1", "2", "3", "4", "5", "6+")
dH2 <- as_tibble(data[data$`Repetition Index` > 0,]) %>% 
   mutate(RepIndex = case_when(
     `Repetition Index` == 1 ~ tags[1],
     `Repetition Index` == 2 ~ tags[2],
     `Repetition Index` == 3 ~ tags[3],
     `Repetition Index` == 4 ~ tags[4],
     `Repetition Index` == 5 ~ tags[5],
     `Repetition Index`  > 5 ~ tags[6]
     ))
dH2$RepIndex <- factor(dH2$RepIndex, levels = tags, ordered = FALSE)
summary(dH2$RepIndex)
```


Box plot of FE for repetition index bins

```{r message=FALSE, warning=FALSE}
pdf(file = "/Users/mario/code/exp-rep/paper/aacl2022/figures/fe-cumul.pdf",
    width = 6,
    height = 6)

ggplot(data = dH2, mapping = aes(x= RepIndex,y=FE)) + 
  geom_jitter(aes(color='blue'),alpha=0.2) +
  geom_boxplot(fill="bisque",color="black",alpha=0.3) + 
  # labs(x='\nRepetition index') + 
  labs(x='\nRepetition Index') +
  labs(y='') +
  guides(color=FALSE) +
  theme_minimal() +
  theme(text = element_text(size=25))

dev.off()
```


### Decay

Group recency values into bins
```{r}
tags <- c("0-30", "30-100", "100-200", "200-400", "400-800", "800+")
d <- data[data$`Repetition Index` > 0,]
dgroup <- as_tibble(d) %>% 
   mutate(Recency = case_when(
   		`RecencyBoth` >= 0 & `RecencyBoth` <= 30 ~ tags[1],
    	`RecencyBoth` > 30 & `RecencyBoth` <= 100 ~ tags[2],
    	`RecencyBoth` > 100 & `RecencyBoth` <= 200 ~ tags[3],
    	`RecencyBoth` > 200 & `RecencyBoth` <= 400 ~ tags[4],
    	`RecencyBoth` > 400 & `RecencyBoth` <= 800 ~ tags[5],
		`RecencyBoth` > 800  ~ tags[6],
	  ))
dgroup$Recency <- factor(dgroup$Recency, levels = tags, ordered = FALSE)
summary(dgroup$Recency)
```


Box plot of FE for the recency bins

```{r message=FALSE, warning=FALSE}
pdf(file = "/Users/mario/code/exp-rep/paper/aacl2022/figures/fe-decay.pdf",
    width = 6,
    height = 6)

ggplot(data = dgroup, mapping = aes(x=Recency,y=FE)) + 
  geom_jitter(aes(color='blue'),alpha=0.2) +
  geom_boxplot(fill="bisque",color="black",alpha=0.3) + 
  # labs(x='\nRecency (tokens)') + 
  labs(x='Distance') +
  labs(y='') +
  guides(color=FALSE) +
  theme_minimal() +
  theme(text = element_text(size=25),
  	axis.text.x = element_text(angle=45, hjust=1))

dev.off()
```


### Construction length

Group constructions by length (number of words)
```{r}
tags = c("3", "4", "5", "6", "7")
dgroup <- as_tibble(data[data$RecencyBoth >= -1,]) %>% 
   mutate(len = case_when(
     Length == 3 ~ tags[1],
     Length == 4 ~ tags[2],
     Length == 5 ~ tags[3],
     Length == 6 ~ tags[4],
     Length == 7 ~ tags[5],
     ))   
dgroup$len <- factor(dgroup$len, levels = tags, ordered = FALSE)
summary(dgroup$len)
```


Box plot of FE for each construction length
```{r message=FALSE, warning=FALSE}
ggplot(data = dgroup, mapping = aes(x=len,y=FE)) + 
  geom_jitter(aes(color='blue'),alpha=0.2) +
  geom_boxplot(fill="bisque",color="black",alpha=0.3) + 
  labs(x='\nConstruction Length') + labs(y='\nFacilitating effect') +
  guides(color=FALSE) +
  theme_minimal()
```


### Repetition index within the utteracce

Group by repetition index within the utterance

```{r}
tags <- c("0", "1", "2", "3", "4", "5+") 
dgroup <- as_tibble(data[data$`Repetition Index` > 0,]) %>% 
   mutate(RepIndexInTurn = case_when(
     IndexInTurn == 0 ~ tags[1],
     IndexInTurn == 1 ~ tags[2],
     IndexInTurn == 2 ~ tags[3],
     IndexInTurn == 3 ~ tags[4],
     IndexInTurn == 4 ~ tags[5],
     IndexInTurn >= 5 ~ tags[6],
     ))

dgroup$RepIndexInTurn <- factor(dgroup$RepIndexInTurn, levels = tags, ordered = FALSE)
summary(dgroup$RepIndexInTurn)
```


Box plot of FE for repetition index in current utterance

```{r message=FALSE, warning=FALSE}

pdf(file = "/Users/mario/code/exp-rep/paper/aacl2022/figures/fe-repindex-in-utt.pdf",
    width = 6,
    height = 6)

ggplot(data = dgroup, mapping = aes(x=RepIndexInTurn,y=FE)) + 
  geom_jitter(aes(color='blue'),alpha=0.2) +
  geom_boxplot(fill="bisque",color="black",alpha=0.3) + 
  labs(x='\nRep. Index in Utterance') + 
  labs(y='\nFacilitating Effect') +
  guides(color=FALSE) +
  theme_minimal() +
  theme(text = element_text(size=25))

dev.off
```


## Referential vs. Non-referential

```{r}
tags <- c("Referential", "Non-referential")

dH1 <- as_tibble(data) %>% 
   mutate(`IsReferential` = case_when(
     `Referential` == TRUE ~ tags[1],
     `Referential` == FALSE  ~ tags[2]
))

dH1$IsTopical <- factor(dH1$IsReferential, levels = tags, ordered = FALSE)
summary(dH1$IsTopical)
```


Referential vs. Non-referential
```{r message=FALSE, warning=FALSE}
ggplot(data = dH1, mapping = aes(x=IsReferential,y=FE)) + 
  geom_jitter(aes(color='blue'),alpha=0.2) +
  geom_boxplot(fill="bisque",color="black",alpha=0.3,range=0.1,outline=FALSE) + 
  labs(x='') + labs(y='Facilitating effect') +
  guides(color=FALSE) +
  theme_minimal()
```


T-tests

```{r}
t.test(FE ~ Referential, data=dH1)
t.test(S ~ Referential, data=dH1)
```



```{r}
firstmentions <- data[data$`RepetitionIndex` == 1,]
repetitions <- data[data$`RepetitionIndex` > 1,]
nrow(firstmentions) / nrow(data)
nrow(repetitions) / nrow(data)

t.test(FE ~ Referential, data=firstmentions)
t.test(S ~ Referential, data=firstmentions)

t.test(FE ~ Referential, data=repetitions)
t.test(S ~ Referential, data=repetitions)
```

### PMI

```{r}
d_generic <- data[data$PMI <= 1,]
d_specific <- data[data$PMI == max(data$PMI),]
summary(d_generic$FE)
summary(d_specific$FE)
```

```{r}
d_generic$group <- 'Agnostic'
d_specific$group <- 'Specific'
plot.data <- rbind(d_generic, d_specific)
```

```{r}
t.test(d_generic$FE, d_specific$FE)
t.test(d_generic$S, d_specific$S)
```


```{r message=FALSE, warning=FALSE}
pdf(file = "/Users/mario/code/exp-rep/paper/aacl2022/submission/figures/fe-pmi.pdf",
    width = 6,
    height = 6)
ggplot(data = plot.data, mapping = aes(x=group,y=FE)) +
  geom_jitter(aes(color='blue'),alpha=0.2) +
  geom_boxplot(fill="bisque",color="black",alpha=0.3) +
  labs(x='\nPMI') +
  labs(y='') +
  guides(color=FALSE) +
  theme_minimal() +
  theme(text = element_text(size=25))
dev.off()
```



### Linear Mixed Effect Models

Only look at repetitions.
```{r}
reps <- data[data$`Repetition Index` > 0,]
```


Most complex FE model; ablation starts from here
```{r}

# fit.fe.all <- lmer(FE ~ 1 + `log Utterance Position` + `log Construction Length` + `log Repetition Index` + `log Distance` + `Previous Same Utterance` + `log Repetition Index in Utterance` + `PMI` + `Referential` + (1 | `Dialogue ID`/Speaker), reps)
# 
# anova(fit.fe.all)

fit.fe.best <- lmer(`Facilitating Effect` ~ 1 + `log Utterance Position` + `log Construction Length` + `log Repetition Index` + `log Distance` + `Previous Same Utterance` + `log Repetition Index in Utterance` + `PMI` + `Referential` + (1 | `Dialogue ID`/Speaker), reps)

summary(fit.fe.best)
```


```{r}
# summary(fit.s.best)
summ(fit.fe.best, scale = TRUE, n.sd = 2,  confint = TRUE, digits = 3)
```

```{r message=FALSE, warning=FALSE}
pdf(file = "/Users/mario/code/exp-rep/paper/aacl2022/figures/lme-coefficients.pdf",
    width = 6,
    height = 6)

plot_summs(fit.fe.best, scale = TRUE, n.sd = 2, plot.distributions = TRUE, colors='#3da4ab')
dev.off
```


```{r}
# fit.s.all <- lmer(S ~ 1 + logCurrentTurn + logLength + logRepetitionIndex + logDistance + PreviousInSameTurn + logRepIndexInTurn + SharedCurrently + SameAsFirstSpeaker + SameAsPreviousSpeaker + logCurrentSpan + PMI + Topical + (1 | `Dialogue ID`/Speaker), reps)

fit.s.best <- lmer(`log Information Content` ~ 1 + `log Utterance Position` + `log Construction Length` + `log Repetition Index` + `log Distance` + `Previous Same Utterance` + `log Repetition Index in Utterance` + `PMI` + `Referential` + (1 | `Dialogue ID`/Speaker), reps)

# summary(fit.s.best)
summ(fit.s.best, scale = TRUE, n.sd = 2,  confint = TRUE, digits = 3)
# summ(fit.s.best, scale = TRUE)
```


```{r message=FALSE, warning=FALSE}
plot_summs(fit.s.best, fit.fe.best, scale = TRUE, n.sd = 2, model.names=c('Information Content', 'Facilitating Effect'))
```

```{r}
props <- read_csv('/Users/mario/code/exp-rep/paper/aacl2022/submission/code/proportions.csv')
fit.prop.all <- lmer(propConstr ~ 1 +  uttID + (1 | dial_id), props)
summ(fit.prop.all, scale = TRUE, n.sd = 2,  confint = TRUE, digits = 3)
```


```{r}
data_non <- read_csv('/Users/mario/code/exp-rep/paper/aacl2022/submission/code/non-constructions.csv')
```
```{r}
t.test(data$S, data_non$S)
t.test(data$FE, data_non$FE)
```


```{r message=FALSE, warning=FALSE}
data$isConstr <- 'Constructions'
data_non$isConstr <- 'Other Sequences'

common_cols <- intersect(colnames(data), colnames(data_non))
plot.data <- rbind(subset(data, select = common_cols), subset(data_non, select = common_cols))

pdf(file = "/Users/mario/code/exp-rep/paper/aacl2022/submission/figures/fe-constr-nonconstr.pdf",
    width = 6,
    height = 6)

ggplot(data = plot.data, mapping = aes(x=isConstr,y=FE)) + 
  geom_jitter(aes(color='blue'),alpha=0.2) +
  geom_boxplot(fill="bisque",color="black",alpha=0.3) + 
  scale_y_continuous(breaks = round(seq(-15, +15, by = 5),1)) +
  labs(x='') + labs(y='Facilitataing Effect') +
  guides(color=FALSE) +
  theme_minimal() +
  theme(text = element_text(size=25)) +
  coord_cartesian(ylim = c(-10, 11.5))

dev.off()
```
